---
title: "Splitting the colleganza social network by date"
author: "JJ Merelo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Splitting the colleganza social network by date}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: ../inst/REFERENCES.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Using data from Puga and Treffler [@serrata], we will, in this vignette, analyze the change of the *colleganza* social network before and after the *Serrata* [@serrata]. This networks includes (extended) Venetian families that engaged in sea trade and were bound by a type of contract called *colleganza*. 
## Set up

The data in Puga and Treffler's [@serrata] paper is sourced from different places, and there is a gap that goes from 1261 to the beginning of the XIV century; this coincides with the Serrata. Let's first load the data

```{r load, warning=FALSE,message=FALSE}
library(serrata.families)
library(igraph)
data("colleganza.graph")
plot(colleganza.graph)
```
This is not too informative, since there are too many nodes; besides, it spans disparate periods. Let's split the network in its two parts

```{r split}
contracts.pre.1261 <- colleganza.slice(to=1261)
plot(contracts.pre.1261)
contracts.post.1310 <- colleganza.slice(from=1310)
plot(contracts.post.1310)
```
There are less contracts after the Serrata, and besides, their connectedness is very different; this probably shows simply the number of contracts, as well as a hint on how the social landscape changed after the Serrata, with many families extinguished and some of them simply devoted to other pursuits.

## Conclusions

A function included in the package allows the splitting of the *colleganza* social network in two parts, before and after the Serrata. Further analysis, like centrality, can be performed on each of those parts.

## References
